Comparing Cross Validation Methods for Penalized Cox Regression

### 2018-04-04 cross-validation.tex
Complete draft, with Updated Figures

\section*{Abstract}
While cross validation is commonly used for selecting tuning parameters in penalized regression, little research has been done to study cross validation methods for penalized Cox regression. Due to its semi-parametric nature, conducting cross validation for Cox models has always been a challenge. We propose two new cross-validation methods for Cox Regression, and compare them to traditional information criteria as well as a cross-validated partial likelihood approach originally proposed by Verweij et al. Our simulation studies show that, in general, cross-validation tends to be conservative (i.e., select smaller models than the ideal choice of tuning parameters) for penalized Cox regression models. However, our proposed approach of cross-validating the linear predictors generally offers the best balance of stability and performance. We also illustrate these approaches on data from studies of gene expression and progression-free survival in cancer patients.

@@

Penalized Cox Regression is an extension of the regular cox regression model. Estimates of the $\beta$ coefficients are obtained by minimizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{m} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta). \end{equation} \citep{Simon2011}. The first part of the objective function is based on the partial likelihood. The second part is a penalty term that depends on some regularization parameters. In this paper, only the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$ was considered. The penalized cox regression can be used for analyzing high dimensional data, where the number of covariates p $\gg$ n. A common application would be to predict the overall survival time of a particular cancer by certain gene expressions.

Selecting the tuning parameter $\lambda$ is crutial to getting LASSO estimations for $\beta$. As $\lambda$ increases, more coefficients would be set to 0 and fewer variables will be selected by the model. When $\lambda$ is too small, the model tends to be redundant. When $\lambda$ is too large, the model would be too conservative to select any variables.

One common approach for selecting the tuning parameter $\lambda$ for linear regression and logistic regression is via cross validation. The idea of K - fold cross validation is illustrated in Figure 1. The data set would be split into K folds. One fold would be treated as the test set and the other K - 1 folds as the training set. The model would first be built on the training set, then fitted to the test set to obtain cross validation error (CVE).  CVE would be calculated for each candidate $\lambda$, then the one that minimizes cross validated error would be selected.

@@

%ZZZZZZZZZZZZZ
\par The most intuitive approach to conduct cross validation in Cox regression is i) fit the model to training set then ii) calculate cross-validated partial likelihood based on the observations from the test set as a measure for the model's predictive accuracy. As will be discussed in Section 3, this approach is unstable when number of events gets small. Verweij elt. \citep{Verweij1993} proposed a stablized version of cross-validated log likehood, which has been widely used and implemented in R packages such as $\tt{glmnet}$. In this paper, we proposed two alternative ways to carry out cross validation. Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the deviance residual \citep{Therneau1990}of the Cox model.

\par We conducted simulation studies to compare how those methods perform in both low dimensional and high dimensional settings with LASSO penalty. We compared the two proposed methods with the cross-validated partial likelihood approaches and traditional information criteria. We showed that the linear predictor approach outperforms other cross validation approaches in various scenarios. We found out that using cross validation to select tuning parameters for LASSO penalized cox regression tends to be a conservative approach in general. Finally, we apply those methods to two high dimensional data sets with time-to-event outcomes.

@@

\par An alternative way to calculate cross-validated log likelihood was proposed by Verweij elt. They defined cross-validated log likelihood of leaving the ith fold out to be \begin{equation}cvl = \sum_{i=1}^{K} \{l(\hat{\beta}_{-i}) - l_{-i}(\hat{\beta}_{-i})\}. \end{equation} When the ith fold is left out, $l(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with all k folds of observations. $l_{-i}(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with observations from the other K-1 folds.

\par We refer to the first definition as ungrouped cross-validated partial likelihood (CVL) and the second definition as grouped cross-validated partial likelihood. Figure 2 illustrates the ideas of the two methods. The grouped CVL uses observations more efficiently than the ungrouped CVL. Since only the left-out fold would be used, the ungrouped cvl is unstable in practice when there are only very few observations in some folder.

@@

Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors. The data would still be split into K folds. Suppose ith fold is left out and the other K-1 folds are used as the training set to get the estimates $\hat{\beta}_{-i}$. Then the cross-validated linear predictors would be calculated based on the observations in the test set:  \begin{equation}\hat{\eta}_{-i} = X_{i}\hat{\beta}_{-i}.\end{equation} After repeating this for all K folds, a whole set of linear predictors  $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...  \hat{\eta}_{-K})$ can be obtained. A partial likelihood can be built over this set of linear predictors: \begin{equation} L(\hat{\eta}_{-}) = \prod_{j=1}^{m} \frac{exp (\hat{\eta}_{-j})}{\sum_{ k \in R(t_{j})}exp (\hat{\eta}_{-k})}\end{equation}.

Cross validated error is defined as the log of $L(\hat{\eta}_{-})$. This idea of cross validating over linear predictors is implemented in the package ncvreg. For all the approaches introduced so far, they are all equivalent to each other in linear regressions. But since they are building over different baselines, they are different in cox regression.

@@

  \subsection{Cross Validated Deviance Residuals}
  %Need to update this section with the new baseline estimation method%
Since the biggest challenge of conducting cross validation for cox regression is due to the missingness of baseline hazard, we propose a second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$. The sum of squares of Martingale residuals is a natural candicate for evaluating the model performance using baseline hazard. Martingale residuals and its normalized form, deviance residuals, were first proposed by \citep{Therneau1990} for Cox regression model diagnostics.

To estimate the baseline hazard, a set of linear predictors $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...  \hat{\eta}_{-K})$ would be obtained in the same way as the cross validated linear predictor approach. $\hat{\eta}_{-i}$ is computed based on $\hat{\beta}_{-i}$s estimated from the training set and observations from the ith test set. With $\hat{\eta}_{-}$, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}. Then for each observations, the Martingale Residual can be calculated: \begin{equation}\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}},\end{equation} where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation. The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation} The sum of squares of the deviance residuals, $\sum_{i}\hat{d}_{i}^2$, would be used as the cross validated error.

@@

  Simulation studies were conducted to compare how those methods behave relative to each other. We generate data with pre-specified baseline hazard, covariate matrix X, coefficient $\beta$ and censoring mechanism. Both low dimensional and high dimentional scenarios were examined. All cross-validation methods mentioned in Section 2 were applied to the data to select the tuning parameter $\lambda$ and produce $\hat{\beta}$ estimates. Cross-validation were also compared to model selection criteria AIC and BIC.

  Survival times were generated from exponential distribution $h(t) = h_{0} exp(X\beta)$, conditioned on covariates. The entries in the covariate matrix $X_{n \times p}$ were independently generated from Normal (0, 1). True coefficients $\beta$  were assumed to have sparsity: $\beta_{p\times 1} = (\beta_{1},\beta_{2}, ..., \beta_{10}, 0, ..., 0)^{T}$.  Censoring status were generated based on binomial distribution.

  Suppose $\hat{\beta}$ is the coefficients estimated by a fitted cox model. We measure the distance between the fitted model and the true model by mean squared error $MSE = E(\hat{\beta} - \beta)^2$. For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$. Then the $\lambda$s that selected by the cross validation or information criteria would be compared to this optimal $\lambda$. If the $\lambda$ chosen by cross validation is smaller than the optimal $\lambda$, then the cross validation method would be considered liberal. If the $\lambda$ chosen by cross validation is larger than the optimal $\lambda$, then the cross validation method would be considered conservative.

    \subsection {Simulations Comparing Cross-validation Approaches}
  %need to change figures to include the new cross-validated deviance residuals
   The simulation experiments were first conducted under lower dimension settings. Number of observations n is set to be 100. Dimension of the data p is set to be 100. 200 replications were used. Censoring percentage, number of folds used in cross-validation, number of non-zero $\beta$s and the magnitude of the $\beta$s were varied in several different simulation scenarios. Patterns and results shown in various simulation scenarios are consistent with one another. Figure 5 and Figure 6 illustrates one simulation scenario, where about 10$\%$ of which were set to be censored and 10 of the $\beta$s were set to be non-zero and 10 folds were used for cross validation. The x - axis of the plots are the magnitude of the non-zero $\beta$ coefficients.

\begin{figure}[h]
    \centering
		\includegraphics[height= 7cm ]{./figures/figure_2.png}
    \caption{Simulation Studies Comparing 4 CV methods in terms of $\lambda$ and MSE ratio. n = 200, p = 1000, magnitude of non-$\beta$s varies from 0.5 to 1.5}
\end{figure}

    $\lambda$s chosen by different cross-validation methods are compared in Figure 5. All values are from taking the mean of 200 replications. The black line represents the optimal $\lambda$. All four cross validation methods consistently select $\lambda$s larger than the optimal one. Since larger $\lambda$ leads larger penalties and fewer variables are selected, all four cross validation methods seem to be conservative in terms of selecting variables. Ungrouped cross-validated log likelihood and the cross-validated linear predictor seem to be more liberal than the other two methods. The cross-validated Deviance residual is the most conservative one and its performance is quite off from the other three methods.

   Mean squared errors are compared in Figure 6. The MSEs of the cross-validation selected models are compared with the MSE of the model given by the optimal $\lambda$. The y-axis represents the ratio of the two. The ungrouped method and the linear predictor approach are outperforming the other methods and are closest to the minimal MSE.

   Simulation studies were also conducted in high dimension scenarios, where the number of observations and dimension of the covariate matrix is more similar to real genetics research. The results are shown in Table 1. Conclusions are close to what is shown in the low dimensional setting. The M- residual approach performs most conservatively. Linear predictor approach and ungrouped approach are relatively more liberal than the other approaches but still slightly more conservative than the optimal $\lambda$.

\begin{table}[h]
\small
\centering
\caption{Simulations Comparing all four Cross-Validation methods, Varying n and p}
\begin{tabular}{llllllllllllllll}
\hline\\[-0.75em]
       & \multicolumn{3}{l}{n = 100, p = 20} &  & \multicolumn{3}{l}{n = 100, p = 100} &  & \multicolumn{3}{l}{n = 100, p = 1000} &  & \multicolumn{3}{l}{n = 500, p = 10000} \\
       & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} \\ \cline{2-4} \cline{6-8} \cline{10-12} \cline{14-16} 
\hline\\[-0.75em]
       &     &        &MSE       &  &   &       & MSE     &    &    &        & MSE      &  &     &        & MSE      \\
       & $\lambda$(SD)     & &Ratio       &  & 	$\lambda$(SD)         & & Ratio     & &	$\lambda$(SD)         & & Ratio      & &	 $\lambda$(SD)         & & Ratio      
\\ 
\cline{2-2} \cline{4-4} \cline{6-6} \cline{8-8} \cline{10-10} \cline{12-12} \cline{14-14} \cline{16-16} 
\\[-0.75em]
Optimal &0.016 (0.009)&&/&&0.042 (0.003)&&/&&0.056 (0.006)&&/&&0.026 (0.001)&&/\\
Grouped &0.035 (0.008)&&2.643&&0.074 (0.009)&&2.236&&0.177 (0.054)&&2.278&&0.053 (0.003)&&77.541\\
Ungrouped &0.030 (0.009)&&2.404&&0.064 (0.010)&&1.801&&0.115 (0.044)&&1.549&&0.041 (0.003)&&56.105\\
LinearPred &0.033 (0.006)&&2.423&&0.066 (0.009)&&1.850&&0.122 (0.039)&&1.673&&0.041 (0.003)&&59.693\\
DevResid &0.099 (0.009)&&13.941&&0.108 (0.009)&&4.110&&0.175 (0.052)&&2.273&&0.075 (0.003)&&76.882\\
\hline
\end{tabular}
\end{table}

 \subsection {Stability}
   While the ungrouped cross-validated likelihood is an intuitive way to carry out cross-validation, it is quite unstable when it runs into cases where the number of observed events in some fold is really small. As is illustrated in Figure 5, cases of undefined cross-validation error would increase when censoring increases or when number of folds increases.

\begin{figure}
    \centering
		\includegraphics[height= 7cm ]{./figures/figure_3.png}
    \caption{Comparing Stability}
\end{figure}

    \subsection {Results of Leave-One-Out Cross Validation}
	%need to change the figures to have fewer lines
    \par We also compared cross-validation with information criteria AIC and BIC. Simulation studies were conducted in lower dimension and higher dimension respectively. As is illustrated in Figure 7, in lower dimension, AIC tends to perform more liberal and BIC tends to perform more conservative than the optimal $\lambda$. While the cross-validation methods are also conservative, most of them are less conservative than BIC. In terms of MSE, cross-validation methods yield smaller MSE than the information criteria. In higher dimension, both BIC and AIC tend to be more liberal than the optimal $\lambda$ and yield larger MSE.

\begin{figure}
    \centering
		\includegraphics[height= 8cm ]{./figures/figure_4.png}
    \caption{LOOCV}
\end{figure}

  \subsection {Simulations on Deviance Residuals}

\begin{table}[h]
	\small
	\caption{Average of Selected $\lambda$ with Different Baseline Estimation for Deviance Residuals}
	\centering
	\begin{tabular}{lllllllll}
	\hline
	\\[-0.75em]
    & Optimal & Linear Pred & KP & KP & Breslow & Breslow & Exact & Exp\\ 
    &  & & & weighted & & weighted & & \\ \cline{2 - 9} 
\textbf{True Baseline: $Exponential(1)$} &&&&&&&&\\
n = 100, p = 10, non-zero $\beta$: 2 & 0.0461 & 0.0665 & 0.1986 & 0.1970 & 0.1908 & 0.1975 & 0.0540 & 0.0539 \\
n = 100, p = 100, non-zero $\beta$: 20 &0.0555 & 0.1841 & 0.2359 & 0.2201 & 0.2146 & 0.2244 & 0.0805 & 0.0985 \\
n = 100, p = 1000, non-zero $\beta$: 20 & 0.0268 & 0.0505 & 0.0948 & 0.0798 & 0.0867 & 0.0809 & 0.0288 & 0.0270 \\
\textbf{True Baseline:} $Weibull$ &&&&&&&& \\
 n = 100, p = 10, non-zero $\beta$: 2 & 0.0476 & 0.0710 & 0.2004 & 0.1969 & 0.1907 & 0.1907 & N/A & 0.2330 \\ 
\hline
\end{tabular}

\end{table}

>>>>

\section*{Abstract}
While cross validation is commonly used for selecting tuning parameters in penalized regression, little research has been conducted to study cross validation methods for penalized Cox regression. Due to its semi-parametric nature, conducting cross validation for Cox models has always been a challenge. We propose two new cross-validation methods for Cox Regression, and compare them to traditional information criteria as well as a cross-validated partial likelihood approach originally proposed by Verweij et al. Our simulation studies show that, in general, cross-validation tends to be conservative (i.e., select smaller models than the ideal choice of tuning parameters) for penalized Cox regression models. However, our proposed approach of cross-validating the linear predictors generally offers the best balance of stability and performance. We also illustrate these approaches on data from studies of gene expression and overall survival in cancer patients.

@@

Penalized Cox Regression is an extension of the regular cox regression model. Estimates of the $\beta$ coefficients are obtained by minimizing the objective function \begin{equation}  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{m} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta). \end{equation} \citep{Simon2011}. The first part of the objective function is based on the partial likelihood. The second part is a penalty term that depends on some regularization parameters. In this paper, only the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$ was considered. The LASSO cox regression can be used to build prediction models for survival outcome based on high dimensional data, where the number of covariates p $\gg$ n. An example of its application would be to make predictions of cancer patients' the overall survival time with their certain gene expressions.

Selecting the tuning parameter $\lambda$ is crutial to getting LASSO estimations for $\beta$. The LASSO penality enforces sparcity to the regression model, so that it carries out an automatic variable selection along the solution path. As $\lambda$ increases, more coefficients would be set to 0 and fewer variables will be selected by the model. When $\lambda$ is too small, the model tends to be redundant. When $\lambda$ is too large, the model would be too conservative to select any variables. In either case, model's predictive accuray would be affected.

One common approach for selecting the tuning parameter $\lambda$ for linear regression and logistic regression is via the K-fold cross validation. The data set would be split into K folds. One fold would be treated as the test set and the other K - 1 folds as the training set. The model would first be built on the training set, then fitted to the test set to obtain cross validation error (CVE).  CVE would be calculated for each candidate $\lambda$, then the one that minimizes cross validated error would be selected.

@@

%ZZZZZZZZZZZZZ
\par Currently, there are two common cross-validation approaches for Cox regression, both having its disadvantages. The most intuitive approach to conduct cross validation in Cox regression is i) fit the model to training set then ii) calculate cross-validated partial likelihood based on the observations from the test set as a measure for the model's predictive accuracy. As will be discussed in Section 2, this approach is unstable when number of events gets small. The second approach was propsoed by Verweij et al. \citep{Verweij1993}. This is a stablized version of cross-validated log likehood, which has been widely used and implemented in R packages such as $\tt{glmnet}$ \citep{glmnet}. Although this approach fixes the stability issue, it tends to yield more conservative models in practice. We hope to improve cross-validation procedure for Cox regression to achieve a good balance between stability and performance.

\par In this paper, we proposed two alternative ways to carry out cross validation. Instead of cross-validating over partial likelihood, we propose to cross validate over the linear predictors of the regression model and cross validate over the deviance residual \citep{Therneau1990}of the Cox model. We conducted simulation studies to compare how those methods perform in both low dimensional and high dimensional settings with LASSO penalty. We compared the two proposed methods with the existing cross-validated partial likelihood approaches. We showed that the linear predictor approach outperforms other cross validation approaches in various scenarios. We found out that using cross validation to select tuning parameters for LASSO penalized cox regression tends to be a conservative approach in general. Finally, we apply those methods to two high dimensional data sets from real studies with time-to-event outcomes.

@@

\par An alternative way to calculate cross-validated log likelihood was proposed by Verweij et al. They defined cross-validated log likelihood of leaving the ith fold out to be \begin{equation}cvl = \sum_{i=1}^{K} \{l(\hat{\beta}_{-i}) - l_{-i}(\hat{\beta}_{-i})\}. \end{equation} When the ith fold is left out, $l(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with all k folds of observations. $l_{-i}(\hat{\beta}_{-i})$ is the log partial likelihood evaluated at $\hat{\beta}_{-i}$ with observations from the other K-1 folds.

\par We refer to the first definition as ungrouped cross-validated partial likelihood (CVL) and the second definition as grouped cross-validated partial likelihood. The upper panels in Figure 2 illustrates the ideas of the two methods. The grouped CVL uses observations more efficiently than the ungrouped CVL. Since only the left-out fold would be used, the ungrouped CVL is unstable in practice when there are only very few observations in some folder.

@@

Besides cross validating over the partial likelihood, an alternative approach is to cross validate over the linear predictors. In the first step, the ith fold would be left out and the other K-1 folds would be used as the training set to get the estimates $\hat{\beta}_{-i}$. Then, the cross-validated linear predictors would be calculated based on the observations in the test set:  \begin{equation}\hat{\eta}_{-i} = X_{i}\hat{\beta}_{-i}.\end{equation} After repeating this for all K folds, a whole set of linear predictors  $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...  \hat{\eta}_{-K})$ can be obtained. A partial likelihood can be built over this set of linear predictors: \begin{equation} L(\hat{\eta}_{-}) = \prod_{j=1}^{m} \frac{exp (\hat{\eta}_{-j})}{\sum_{ k \in R(t_{j})}exp (\hat{\eta}_{-k})}\end{equation}.

Cross-validated error is defined as the log of $L(\hat{\eta}_{-})$. This idea of cross validating over linear predictors is implemented in the package \texttt{ncvreg} \citep{ncvreg}. For all the approaches introduced so far, they are all equivalent to each other in linear regressions. But since they are building over different baselines, they are different in cox regression.

@@

  \subsection{Cross Validated Deviance Residuals}
  %Need to update this section with the new baseline estimation method%
The fundamental challenge of conducting cross validation for cox regression is that the baseline hazard is not estimated from the model. Hence we proposed our second approach which involves estimating the actual baseline hazard $\hat{\Lambda}_{0}$. The sum of squares of Martingale residuals is a natural candicate for evaluating the model performance using baseline hazard. Martingale residuals and its normalized form, deviance residuals, were first proposed by \citep{Therneau1990} for Cox regression model diagnostics.

To estimate the baseline hazard, a set of linear predictors $\hat{\eta}_{-} = ( \hat{\eta}_{-1},  \hat{\eta}_{-2} , ...  \hat{\eta}_{-K})$ would be obtained in the same way as the cross validated linear predictor approach. $\hat{\eta}_{-i}$ is computed based on $\hat{\beta}_{-i}$s estimated from the training set and observations from the ith test set. With $\hat{\eta}_{-}$, a baseline hazard $\hat{\Lambda}_{0}$ can be estimated via Kalbfleisch and Prentice's method \citep{Kalbfleisch2011}. Then for each observations, the Martingale Residual can be calculated: \begin{equation}\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}_{i}},\end{equation} where $\delta_{j}$ is the status of the jth observation and $t_{j}$ is the time component of the jth observation. The deviance residual can be calcualted from the Martingale Residuals: \begin{equation} d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}\end{equation} The sum of squares of the deviance residuals, $\sum_{i}\hat{d}_{i}^2$, would be used as the cross validated error. An illustration of this method is drawn in the bottom right panel of Figure 1.

@@

  Simulation studies were conducted to compare how those methods behave relative to each other. We generated data with pre-specified baseline hazard, covariate matrix X, coefficient $\beta$ and censoring mechanism. Both low dimensional and high dimentional scenarios were examined. All cross-validation methods mentioned in Section 2 were applied to the data to select the tuning parameter $\lambda$ and produce $\hat{\beta}$ estimates. %Cross-validation were also compared to model selection criteria AIC and BIC.

 The entries in the covariate matrix $X_{n \times p}$ were independently generated from $Normal(0, 1)$. True coefficients $\beta$ were assumed to have sparsity: $\beta_{p\times 1} = (\beta_{1},\beta_{2}, ..., \beta_{j}, 0, ..., 0)^{T}$. The covariate matrix $X_{n \times p}$ was centered before it was used to generate outcomes. Survival times were generated from exponential distribution $h(t) = h_{0} exp(X\beta)$, conditioned on covariates. Censoring status were generated based on binomial distribution. All simulations were implemented in R \citep{R}.

  Suppose $\hat{\beta}$ is the coefficients estimated by a fitted cox model. We quantify the predictive accuracy by measuring the distance between the fitted model and the true model by mean squared error MSE $= E(\hat{\beta} - \beta)^2$. For each generated data set, the $\lambda$ that has the minimal MSE is chosen as the optimal $\lambda$. Then the $\lambda$s that selected by the cross validation or information criteria would be compared to this optimal $\lambda$. If the $\lambda$ chosen by cross validation is smaller than the optimal $\lambda$, then the cross validation method would be considered liberal. If the $\lambda$ chosen by cross validation is larger than the optimal $\lambda$, then the cross validation method would be considered conservative.

    \subsection {Simulations Comparing Cross-validation Approaches}
  %need to change figures to include the new cross-validated deviance residuals
   The simulation experiments were first conducted when the signal in the data was varied. In each simulated data set, there are n = 100 observations and p = 1000 covariates. 10$\%$ of the observations were censored. 10-fold cross-validation was implemented for all four methods. Number of non-zero $\beta$s was 20. The signal in the data set was varied by varing the magnitude of the $\beta$s from 0.50 to 1.50. For each scenario, 200 replications were used.

\begin{figure}[h]
    \centering
		\includegraphics[height= 7cm ]{./figures/figure_2.png}
    \caption{Simulation Studies Comparing 4 CV methods in terms of $\lambda$ and MSE ratio. n = 200, p = 1000, 10$\%$ censoring.}
\end{figure}

\par Results of this simulation were illustrated in Figure 2.  The x axis of the plots are the magnitude of the non-zero $\beta$ coefficients. The left panel compares the $\lambda$ selected by different methods. The grouped cross-validated partial likelihood approach, which is currently most widely used approach, consistently selects $\lambda$s that are greater than the optimal one.  Greater $\lambda$ leads to greater penalties and fewer non-zero variables, hence the grouped approach is conservative in terms of selecting variables. While the proposed cross-validated Deviance residual has very close performance as the grouped approach, the ungrouped cross-validated log likelihood and the cross-validated linear predictor were more liberal than the other two methods. In general, all four methods tends to act conservatively and selects smaller models.

\par The right panel compares the MSE ratios of the model selected. The MSEs of the cross-validation selected models are compared with the MSE of the model given by the optimal $\lambda$. The y-axis represents the ratio of the two. The ungrouped method has the smallest MSE ratio. The linear predictor approach's performance is also very close the ungrouped appraoch. Both the grouped approach and the proposed deviance residual approach yield models that have greater MSE, hence are not recommended for use in practice.

\begin{table}[h]
\small
\centering
\caption{Simulations Comparing all four Cross-Validation methods, Varying n and p}
\begin{tabular}{llllllllllllllll}
\hline\\[-0.75em]
       & \multicolumn{3}{l}{n = 100, p = 20} &  & \multicolumn{3}{l}{n = 100, p = 100} &  & \multicolumn{3}{l}{n = 500, p = 10000} &  & \multicolumn{3}{l}{n = 250, p = 10000} \\
       & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 10} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} &  & \multicolumn{3}{l}{$\#$of non-zero = 20} \\ \cline{2-4} \cline{6-8} \cline{10-12} \cline{14-16} 
\hline\\[-0.75em]
       &     &        &MSE       &  &   &       & MSE     &    &    &        & MSE      &  &     &        & MSE      \\
       & $\lambda$ (SD)     & &Ratio       &  & 	$\lambda$ (SD)         & & Ratio     & &	$\lambda$ (SD)         & & Ratio      & &	 $\lambda$ (SD)         & & Ratio      
\\ 
\cline{2-2} \cline{4-4} \cline{6-6} \cline{8-8} \cline{10-10} \cline{12-12} \cline{14-14} \cline{16-16} 
\\[-0.75em]
Optimal &0.016 (0.009)&&/&&0.042 (0.003)&&/&&0.026 (0.001)&&/&&0.026 (0.001)&&/\\          
Grouped &0.035 (0.007)&&2.597&&0.073 (0.009)&&2.184&&0.053 (0.003)&&5.627&&0.123 (0.036)&&3.931\\
Ungrouped &0.030 (0.009)&&2.246&&0.064 (0.01)&&1.815&&0.040 (0.003)&&3.131&&0.070 (0.016)&&2.560\\
LinearPred &0.034 (0.007)&&2.559&&0.066 (0.008)&&1.880&&0.041 (0.003)&&3.350&&0.074 (0.016)&&2.713\\
DevResid &0.100 (0.009)&&14.005&&0.108 (0.008)&&4.087&&0.076 (0.003)&&9.133&&0.111 (0.028)&&3.767\\ 
\hline
\end{tabular}
\end{table}

\par Simulation studies were also conducted in scenarios with various sample sizes and dimensions. The results are shown in Table 1. While the ungrouped approach yields the smallest MSE and best predictive accuracy, the Linear Predictor and the Ungrouped approaches perform closely to each other across all settings. Grouped approach performs well in low-dimensions or with sufficient samples, while deviance residual approach performs better in high-dimensional settings.

 \subsection {Stability}

\par In this section, simulations studies were conducted to address the stability issue of the ungrouped cross-validated likelihood. The ungrouped approach is an intuitive way to carry out cross-validation. It also has the best predictive accuarcy based on the simulation results in previous sections. However, one big disadvantage of this approach is that it is unstable when number of observed events in some fold is really small. As is illustrated in Figure 3, when censoring increases or when number of folds increases, the model would fail to converge due to insufficient number of events in the test set. Hence, the ungrouped cross-validated likelihood should not be used when there is no sufficient number of events.

\begin{figure}[h]
    \centering
		\includegraphics[height= 7cm ]{./figures/figure_3.png}
    \caption{Comparing Stability}
\end{figure}

    \subsection {Results of Leave-One-Out Cross Validation}
	\par In this section, Leave-One-Out Cross Validation (LOOCV) were studied, in comparison with 10-fold CV. When first introduced, the grouped cross-validated partial likelihood was proposed so that LOOCV for Cox regression became feasible. It is impossible to carry out LOO cross-validation using the ungrouped approach, because the partial likelihood cannot be built on the one left-out observation. While the grouped cross-validated partial likelihood solves the stability issue by aggregating more observations together, it is not an efficient way to evaluate the model fit. Simulation studies in Section 3.1 have illustrated that it tends to be conservative in variable selections and selects smaller models in various scenarios. This fact can be even more illustrated via the following simulation studies, where both 10-fold and LOOCV were conducted via grouped, linear predictor and deviance residual approaches and 10 fold CV was conducted for the ungrouped approach.
	\par As is illustrated in Figure 4, LOOCV selected models that have smaller MSE than 10-fold CV for all three CV approaches. As is shown in left and middle panels, LOOCV based on deviance residuals and grouped cross-validated partial likelihood still perform more conservatively than the ungrouped approach. Even if the grouped approach was proposed to allow LOOCV, the LOOCV based on the grouped approach does not surpass the 10-fold ungrouped approach. On the other hand, the Linear Predictors also allows LOOCV, and the LOOCV based on Linear Predictors performs better than the ungrouped approach.

    %\par We also compared cross-validation with information criteria AIC and BIC. Simulation studies were conducted in lower dimension and higher dimension respectively. As is illustrated in Figure 7, in lower dimension, AIC tends to perform more liberal and BIC tends to perform more conservative than the optimal $\lambda$. While the cross-validation methods are also conservative, most of them are less conservative than BIC. In terms of MSE, cross-validation methods yield smaller MSE than the information criteria. In higher dimension, both BIC and AIC tend to be more liberal than the optimal $\lambda$ and yield larger MSE.

\begin{figure}[h]
    \centering
		\includegraphics[height= 8cm ]{./figures/figure_4.png}
    \caption{LOOCV}
\end{figure}

  \subsection {Simulations on Deviance Residuals and Baseline Hazard Estimations}
\par While the Deviance Residual approach has a reasonable performance in scenarios where dimension is high, it tends to act more conservatively than all the other three approaches in general. As is noted by Therneau et al, the sums of squared martingale residuals and deviance residuals do not necessary reflect how good the model fit is \citep{Therneau2000modeling}. The difficulty of estimating the baseline hazard seems to play an important role. A series of simulations were thus designed to study how baseline hazard estimation affects the deviance residuals' performance as a measurement for model fit.

\par In the following simulation studies, the defined optimal $\lambda$ and the $\lambda$ selected by the Linear Predictor approach were used as benchmarks. Various sample sizes and dimensions were varied. The survival outcomes were generated both from exponential distribution and Weibull distribution. Six different approaches were compared in terms of baseline hazard estimation and their impact on sums of squared deviance residuals. Two non-parametric approaches, Kalbfleisch and Prentices' estimator and Breslow's estimator, were used first for estimating the cumulative hazard baseline \citep{Kalbfleisch2011} \citep{breslow1972}.  As is presented in column 3 and column 5 in Table 2, the sum of squared deviance residuals computed based on both approaches are all very conservative compared to the optimal $\lambda$. Their performance is extremely bad when n is large and p is small.

\par Since both of the non-parameteric estimations seemed to perform well at early time points but poorly later, a weighted approach was then applied to emphasize earlier time points. Weights are assigned proportional to number of subjects at risk, making earlier time points weighted more than later time points. Results of these two approahces are presented in column 4 and 6 in Table 2. The attempt to improve the non-parameteric estimations via weighting does not seem to improve the selection too much.

\par In the next step,  the non-parametric baseline hazard estimation was replaced by the exact baseline hazard that was used to generate the outcome. Results were listed in the seventh column in Table 2. When the baseline was known, the performance of sum of squared deviance residuals actually performed really close to the optimal $\lambda$. This simulation suggests that when the baseline hazard was accurately estimated, then the sum of squared deviance residuals gave really good measures for the model fit.

\par In the last step, we replaced the non-parametric baseline hazard estimation with a parametric approach, where we assume the baseline hazard to follow an exponential distribution. We derived the estimation for baseline hazard by maximum likelihood approach conditioned on covariates: $\hat{\lambda}_{0} = \frac{\sum d_{i}}{\sum t_{i}e^{\eta_{i}}}$. The results are listed in the eighth column in Table 2. When this estimation was used for data that were generated from exponential distribution, its performance was as well as when the exact baseline was specified. When it was fitted to data that were generated from the Weibull distribution, that is when the model is misspecified, then its performance was even worse than the non-parametric approaches.

\par Hence, getting accurate estimation of the baseline hazard is crucial to whether or not the sum of squared deviance residuals can be a good measurement for model fit.
